---
title: "arimamax_tsfinal"
author: "ANambiar"
date: "2023-12-01"
output: html_document
---
```{r setup, include=FALSE}
library(readr)
library(tseries)
library(fpp)
library(ggplot2)
library(forecast)
library(janitor)
library(dplyr)
library(lubridate)
library(TSA)
library(MuMIn)
```

### edited data loading and manipulating

```{r Loading the dataset}
data <- read_csv("/Users/pandita/Documents/merged.csv", show_col_types = FALSE)
data <- data %>%
  mutate(across(where(is.numeric), ~ifelse(. == "Not Available", 0, .)))
data$Date <- as.Date(data$Date, format = "%Y-%m-%d")  # Adjust format as per your date format
df <- clean_names(data)

df_pre_2005 <- df %>% filter(date < as.Date("2005-08-01"))
df_post_2005 <- df %>% filter(date >= as.Date("2005-08-01") & date <= as.Date("2009-03-01"))
```

```{r}
# Function to create time series objects
create_time_series_objects <- function(df, start_date) {
  ts_list <- list()
  for (col_name in colnames(df)) {
    if (col_name != "date") {  
      col_data <- df[[col_name]]
      ts_obj <- ts(col_data, start = start_date, frequency = 12)
      ts_name <- paste0("ts_", col_name)
      ts_list[[ts_name]] <- ts_obj
    }
  }
  return(ts_list)
}

# Create time series objects for pre-2005 and post-2005 datasets
ts_pre_2005 <- create_time_series_objects(df_pre_2005, c(1973, 1, 1))
ts_post_2005 <- create_time_series_objects(df_post_2005, c(2005, 8, 1)) 

pre_2005_train <- ts_pre_2005$ts_total_energy_co2_emissions
post_2005_train <- ts_post_2005$ts_total_energy_co2_emissions
```

### stationary check 

```{r Checking for stationarity}
# Check for stationarity
adf.test(pre_2005_train, alternative="stationary")
# Since the data is non-stationary, apply differencing to detrend the data (why in other notebook) 
seasonal_diff <- diff(pre_2005_train, lag=12, differences=1)
```

###  arimax

### figure out p,q for non-seasonal components from EACF plot

```{r}
eacf(seasonal_diff)
```

Non-Seasonal Differencing (d) = 1
Seasonal Differencing (D) = 1
ARIMA(p, 0, q)(P, 1, Q)

Lowest p + q combos:
ARIMA(1, 1, 3), ARIMA(3, 1, 0), ARIMA(2, 1, 2) 

use values around (2,1) for seasonal components

```{r}
non_seasonal_orders <- list(c(1, 3), c(3, 2), c(2, 2))
seasonal_order_range <- list(P = c(1, 2, 3), Q = c(0, 1, 2))
```

### arimax function

```{r}
fit_arimax_models <- function(data, xreg, non_seasonal_orders, seasonal_order_range) {
  results <- list()
  best_model <- NULL
  lowest_aicc <- Inf
  
  # Fit ARIMAX models with specified orders
  for(order in non_seasonal_orders) {
    for(P in (seasonal_order_range$P[1]):(seasonal_order_range$P[3])) {
      for(Q in (seasonal_order_range$Q[1]):(seasonal_order_range$Q[3])) {
        current_order <- c(order[1], 1, order[2])
        current_seasonal_order <- c(P, 1, Q)
        
        model_attempt <- try(Arima(data, order = current_order, seasonal = current_seasonal_order, xreg = xreg), silent = FALSE)
        
        if (inherits(model_attempt, "try-error")) {
          print(paste("Error in fitting model:", current_order, current_seasonal_order))
          print(model_attempt)
        } else {
          model <- model_attempt
          model_name <- paste("ARIMA", paste(current_order, collapse = ","), " x ", paste(current_seasonal_order, collapse = ","), sep = "")
          aicc_value <- AICc(model)
          results[[model_name]] <- list(Model = model, AICc = aicc_value)
          cat(model_name, ": AICc =", aicc_value, "\n")
          
          if (aicc_value < lowest_aicc) {
            best_model <- model
            lowest_aicc <- aicc_value
          }
        }
      }
    }
  }

  # Fit auto.arima model with regression
  auto_model_attempt <- try(auto.arima(data, xreg = xreg, seasonal = TRUE), silent = FALSE)
  
  if (inherits(auto_model_attempt, "try-error")) {
    print("Error in fitting auto.arima model")
    print(auto_model_attempt)
  } else {
    auto_model <- auto_model_attempt
    arima_order <- auto_model$order
    seasonal_order <- auto_model$seasonal$order
    frequency <- auto_model$seasonal$period
    aicc_value <- AICc(auto_model)
    results[["auto.arima"]] <- list(Model = auto_model, AICc = aicc_value)
    cat("auto.arima: AICc =", aicc_value)
    
    if (aicc_value < lowest_aicc) {
      best_model <- auto_model
      lowest_aicc <- aicc_value
    }
  }
    results[["best_model"]] <- best_model
  return(results)
}
```

### xreg for pre 2005

```{r}
# filtering columns
#pre_2005_df_filtered <- df_pre_2005 %>% select(-contains("co2_emissions"), -date)
pre_2005_df_filtered <- df_pre_2005 %>% 
  select('gdp', 'total_renewable_energy_consumption', 'total_renewable_energy_production', 'total_fossil_fuels_production', 'total_fossil_fuels_consumption')

# make numeric, deal with nulls
xreg <- pre_2005_df_filtered
xreg <- data.frame(lapply(xreg, function(x) as.numeric(as.character(x))))
#colSums(is.na(xreg))
#xreg <- xreg[, colSums(is.na(xreg)) == 0]
if (!is.matrix(xreg)) {
  xreg <- as.matrix(xreg)
}
```


```{r}
auto.arima(pre_2005_train, xreg = xreg, seasonal = TRUE)
```


### models for pre 2005

```{r}
fitted_models <- fit_arimax_models(pre_2005_train, xreg, non_seasonal_orders, seasonal_order_range)
```


Best model: ARIMA2,1,2 x 1,1,2 : AICc = 1414.293 


### xreg for post 2005

```{r}
# filtering columns
post_2005_df_filtered <- df_post_2005 %>% 
  select('gdp', 'total_renewable_energy_consumption', 'total_renewable_energy_production', 'total_fossil_fuels_production', 'total_fossil_fuels_consumption')

# make numeric, deal with nulls
xreg_post <- post_2005_df_filtered
if (!is.matrix(xreg_post)) {
  xreg_post <- as.matrix(xreg_post)
}
```

### models for post 2005

```{r}
fitted_models_post <- fit_arimax_models(post_2005_train, xreg_post, non_seasonal_orders, seasonal_order_range)
```

Best model: auto.arima: AICc = 154.9541
### pre 2005: checkresiduals for best model, ljung box for residuals of best model - make sure white noise

```{r}
best_model_pre2005 <- fitted_models[["best_model"]]
checkresiduals(best_model_pre2005)
ljung_box_test <- Box.test(best_model_pre2005$residuals, lag = 20, type = "Ljung-Box")
print(ljung_box_test)
```

```{r}
auto.arima(post_2005_train, xreg = xreg_post, seasonal = TRUE)
```


### post 2005: checkresiduals for best model, ljung box for residuals of best model - make sure white noise

```{r}
best_model_post2005 <- fitted_models_post[["best_model"]]
checkresiduals(best_model_post2005)
ljung_box_test <- Box.test(best_model_post2005$residuals, lag = 20, type = "Ljung-Box")
print(ljung_box_test)
```
```{r}
# forecast best model after 2005 July policy implementation
forecast_best_model_2005 <- forecast(best_model_pre2005, xreg = xreg_post)
```


```{r}
# plot forecasted value with the actual values
# Plot the forecast
plot(forecast_best_model_2005, main = "Forecast vs Actual", xlab = "Time", ylab = "Values")

# Add the actual values
lines(post_2005_train, col = "red")

# Add a legend
legend("topleft", legend = c("Forecast", "Actual"), col = c("blue", "red"), lty = 1)

```

```{r}
# comparing best models pre policy
best_model_pre2005
```


```{r}
# comparing best models pre policy
best_model_post2005
```
