---
title: "TS_Project_Puneet"
author: "Puneet Modi"
date: "2023-11-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
library(readr)
library(tseries)
library(fpp)
library(ggplot2)
library(forecast)
library(janitor)
library(dplyr)
library(lubridate)
library(TSA)
library(MuMIn)
```


```{r Loading the dataset}
data <- read_csv("merged.csv")

# Imputing missing values
data <- data %>%
  mutate(across(where(is.numeric), ~ifelse(. == "Not Available", 0, .)))

# Converting date to a datetime object
data$Date <- as.Date(data$Date)

# Cleaning up the column names
df <- clean_names(data)

head(df)

# Creating time series objects for all columns
create_time_series_objects <- function(df) {
  # Initialize an empty list to store the time series objects
  ts_list <- list()
  
  # Iterate through the columns of the data frame
  for (col_name in colnames(df)) {
    # Extract the column data
    col_data <- df[[col_name]]
    
    # Create a time series object with the specified start and frequency
    ts_obj <- ts(col_data, start = c(1973, 1), frequency=12)
    
    # Name the time series object
    ts_name <- paste0("ts_", col_name)
    
    # Assign the time series object to the list with the specified name
    ts_list[[ts_name]] <- ts_obj
  }
  
  # Return the list of time series objects
  return(ts_list)
}

# Creating a list that contains ts objects for every column 
ts <- create_time_series_objects(df)
```


```{r Train and Validation Sets}
# Train and Validation sets are pre-2005. Forecasting post-2005 to show the 
# effect of the policy on CO2 Emissions
train_data_2005 <- window(ts$ts_total_energy_co2_emissions, start=1973, end=2005)
train_data_2009 <- window(ts$ts_total_energy_co2_emissions, start=1973, end=2009)
# validation_data <- window(ts$ts_total_energy_co2_emissions, start=1973, end=2005)

# Plotting the dependent variable -> total CO2 emissions
plot(ts$ts_total_energy_co2_emissions, main="Total CO2 Emissions")
acf(ts$ts_total_energy_co2_emissions)
pacf(ts$ts_total_energy_co2_emissions)

# Applying Box-Cox transformation to stabilize the variance 
lambda <- BoxCox.lambda(ts$ts_total_energy_co2_emissions)
lambda # 1.41

# Separating the seasonal component
decomposed_data <- stl(ts$ts_total_energy_co2_emissions, s.window="periodic")
plot(decomposed_data)
```

We observe two trends - an increasing trend till 2005 and a decreasing trend post 
2005. Hence, total CO2 emissions is not stationary. 

We observed that as time increases, the amplitude seems to increase and decrease. 
This change in the height of the peaks/troughs does not seem to be constant over
time. The data shows multiplicative seasonality.


```{r Checking for stationarity}
# Check for stationarity
adf.test(train_data_2005, alternative="stationary")

# Since the data is non-stationary, apply differencing to detrend the data and 
# make it stationary. Second order leads to over differencing. 
seasonal_diff <- diff(train_data_2005, lag=12, differences=1)
seasonal_diff_2 <- diff(train_data_2005, lag=12, differences=2)
plot(seasonal_diff)

# ACF for seasonal differencing
acf(seasonal_diff, main="ACF of Seasonally Differenced Series")

# PACF for seasonal differencing 
pacf(seasonal_diff, main="PACF of Seasonally Differenced Series")

# KPSS test on original data
kpss_test_original <- kpss.test(train_data_2005)
print(paste("KPSS Test for Original Data: ", kpss_test_original$p.value))

# KPSS test on 1st order difference
kpss_test_diff1 <- kpss.test(seasonal_diff)
print(paste("KPSS Test for 1st Order Difference: ", kpss_test_diff1$p.value))
```

According to the ADF test, the data is not stationary. The p-value is greater 
than 0.05 so we accept the null hypothesis. We then find an appropriate 
differencing which yields seasonal and trend stationary data. 

Plot the ACF and PACF to determine if the detrended and deseasonalized time 
series is stationary.

```{r Creating labels for intervention analysis}
# Finding the best model pre-2005
model_2005 <- auto.arima(seasonal_diff, seasonal=TRUE, 
                         stationary=TRUE, lambda=lambda) #############
model_2005
checkresiduals((model_2005))

eacf(seasonal_diff)

# Based on eacf, fitting different ARIMA models and comparing the AICc and 
# BIC values
fit1_2005 <- Arima(train_data_2005, order=c(1, 0, 0), seasonal=c(2, 1, 1))
summary(fit1_2005)
fit2_2005 <- Arima(train_data_2005, order=c(1, 0, 1), seasonal=c(2, 1, 1))
summary(fit2_2005)
fit3_2005 <- Arima(train_data_2005, order=c(2, 0, 1), seasonal=c(2, 1, 1))
summary(fit3_2005)
fit4_2005 <- Arima(train_data_2005, order=c(2, 0, 2), seasonal=c(2, 1, 1))
summary(fit4_2005)

# Finding the best model pre-2009
model_2009 <- auto.arima(train_data_2009, seasonal=TRUE, 
                         stationary=FALSE, lambda=lambda) #############
model_2009
checkresiduals((model_2009))

eacf(train_data_2009[1:432])

# Based on eacf, fitting different ARIMA models and comparing the AICc and 
# BIC values
fit1_2009 <- Arima(train_data_2009, order=c(1, 0, 0), seasonal=c(2, 1, 1))
summary(fit1_2009)
fit2_2009 <- Arima(train_data_2009, order=c(1, 0, 1), seasonal=c(2, 1, 1))
summary(fit2_2009)
fit3_2009 <- Arima(train_data_2009, order=c(2, 0, 1), seasonal=c(2, 1, 1))
summary(fit3_2009)
# fit4_2009 <- Arima(train_data_2009, order=c(2, 0, 2), seasonal=c(2, 1, 1))
# summary(fit4_2009)

# # Exponential smoothing model using ets()
# model_ets <- ets(ts$ts_total_energy_co2_emissions, damped="Multiplicative", 
#                  allow.multiplicative.trend=TRUE, lambda="auto")
# summary(model_ets)

# df$Intervention <- ifelse(year(data$Date) >= 2005, 1, 0)
# 
# model <- arimax(ts$ts_total_energy_co2_emissions, order=c(1,0,2), xreg=df$Intervention)
# model
```